Supplementary Materials: Introductory Physics Labs: A Tale of Two Transformations

### 2020-09-26 manuscript/onlineLabPaper.tex
Updated manuscript draft, and edited the supplement.  Needs more work, but should be ready for review in the next few days.

\section{Introduction}
A significant challenge facing physics faculty face teaching introductory labs is engaging
students in authentic science practices.  Another has been highlighted given the current global
pandemic---how to engage students in our laboratory courses while maintaining appropriate
social distancing and hygiene standards. We have chosen to answer these challenges by
transforming our labs\ldots\ twice. We discuss the rationale behind the first transformation to
a practice-focused curriculum.  In March 2020 we needed to transform our labs again, this time
to accommodate online learning. This paper discusses two chief questions: ``What are we doing
to engage students in science practices?'' and ``How did we make all of this work online?''

@@

The physics labs at our institution in Fall 2016 very much mirrored the national narrative.
The lab curriculum had been largely unchanged for about 20 years. Moreover, the equipment and techniques used by students was outdated---triple beam balances and hand-drawn graphs were
standard.  That same year, as part of an interdisciplinary team interested in engaging
students in and assessing scientific practices, we secured
funding, both internally and from the NSF, to change this \emph{status quo} in physics, biology, and
chemistry.  All three departments transformed their labs using the Argument Driven Inquiry (ADI),\citep{Sampson2011,Walker2011,Walker2016} pedagogical framework, which has
been specifically designed to engage students in authentic science practices. In physics, we
also utilized the AAPT Lab Recommendations to guide our transformation.\citep{kozminski2014aapt}

@@

Regional State University is a public doctoral granting university in a rural, economically
depressed part of our state.  The university mission is focused on regional transformation and
service to this region.  The total combined student population in Fall 2019 was 28,651 largely
comprised of undergraduates (23,081). Our physics labs (Physics \textrm{I} lab and Physics
\textrm{II} lab) are each 1 credit courses that meet for 2 hours per week. These courses serve
both of our Calculus-based and Algebra-based physics lecture courses.  Most $(\sim 75\%)$ of
the students who take our lab course are in (or have taken) the corresponding Algebra-based
physics lecture course. These labs also fulfill the general education requirement for science
lab courses at our university.

@@

The argumentation session was held in a scheduled Cisco WebEx session
the week after the proposal session. Before this session, each group was required to complete
its analysis and create a three-slide presentation for the argumentation session: (1) a description of their measurements, (2) a presentation of the
results, including a graph or table, and (3) their argument based on their
results.  One member of each group gave the presentation, which was followed by
questions. Students received credit for giving presentations, asking meaningful questions, and
responding to questions.  Following the argumentation session students submitted individual
draft reports, peer-reviewed each other's drafts, and submitted final reports in the same
manner used for the face-to-face investigations.

@@

\section{Discussion}
We have observed a great number of successes for these transformations.  First and foremost, we indeed have a lab curriculum that is aligned with national goals and
standards. Furthermore, we have successfully deployed this online with both short and longer
planning times.  Students are engaging with science practices as determined by our
practical assessments, and our department culture around laboratory course instruction is
invigorated. Psychometric assessment of our students engagement with science practices is
ongoing.

>>>>

\section{Introduction}
A significant challenge facing physics faculty face teaching introductory labs is engaging
students in authentic science practices.\citep{national2007Rising,PCAST12,kozminski2014aapt}
Another has been highlighted given the current global pandemic---how to engage students in our
laboratory courses while maintaining appropriate social distancing and hygiene standards. We
have chosen to answer these challenges by transforming our labs\ldots\ twice. We discuss the
rationale behind the first transformation to a practice-focused curriculum.  In March 2020 we
needed to transform our labs again, this time to accommodate online learning. This paper
discusses two chief questions: ``What are we doing to engage students in science practices?''
and ``How did we make all of this work online?''

@@

The physics labs at our institution in Fall 2016 very much mirrored the national narrative.
The lab curriculum had been largely unchanged for about 20 years. Moreover, the equipment and
techniques used by students was outdated---triple beam balances and hand-drawn graphs were
standard.  That same year, as part of an interdisciplinary team interested in engaging students
in and assessing scientific practices, we secured funding, both internally and from the NSF, to
change this \emph{status quo} in physics, biology, and chemistry.  All three departments
transformed their labs using the Argument Driven Inquiry
(ADI)\citep{Sampson2011,Walker2011,Walker2016} pedagogical framework, which has been
specifically designed to engage students in authentic science practices. In physics, we also
utilized the AAPT Lab Recommendations to guide our transformation.\citep{kozminski2014aapt}

@@

% Regional State University is a public doctoral granting university in a rural, economically
% depressed part of our state.  The university mission is focused on regional transformation and
% service to this region.  The total combined student population in Fall 2019 was 28,651 largely
% comprised of undergraduates (23,081). Our physics labs (Physics \textrm{I} lab and Physics
% \textrm{II} lab) are each 1 credit courses that meet for 2 hours per week. These courses serve
% both of our Calculus-based and Algebra-based physics lecture courses.  Most $(\sim 75\%)$ of
% the students who take our lab course are in (or have taken) the corresponding Algebra-based
% physics lecture course. These labs also fulfill the general education requirement for science
% lab courses at our university.

@@

The argumentation session was held in a scheduled Cisco WebEx session
the week after the proposal session. Before this session, each group was required to complete
its analysis and create a three-slide presentation for the argumentation session: (1) a description of their measurements, (2) a presentation of the
results, including a graph or table, and (3) their argument based on their
results.  One member of each group gave the presentation, which was followed by
questions. Students received credit for giving presentations, asking meaningful questions, and
responding to questions.  Following the argumentation session, students submitted individual
draft reports, peer-reviewed each other's drafts, and submitted final reports in the same
manner used for the face-to-face investigations.

@@

\section{Discussion}
We have observed a great number of successes for these transformations.  First and foremost, we
indeed have a lab curriculum that is aligned with national goals and standards. Furthermore, we
have successfully deployed this online.  Students are engaging with science practices as
determined by our practical assessments, and our department culture around laboratory course
instruction is invigorated. Psychometric assessment of our students engagement with science
practices is ongoing.

### 2020-09-26 supplement/supplement.tex
Updated manuscript draft, and edited the supplement.  Needs more work, but should be ready for review in the next few days.

\section{Face-to-face implementation}

Activities and basic equipment.

Assessment practices and discussion of results.

\section{Online implementation}

Curricular adaptation.

Kit contents.

\section{Introduction}
There are many challenges that physics faculty face when teaching introductory labs. One of
these challenges is engaging students in authentic science practices.  Another has been
highlighted given the current global pandemic---specifically how to engage students in our
laboratory courses while maintaining appropriate social distancing and hygiene standards. We
have chosen to answer these challenges by transforming our labs\ldots\ twice. We discuss the
rationale behind the first transformation to a practice-focused curriculum.  However, when
March 2020 rolled around we needed to transform our labs again, this time to accommodate online
learning. This paper discusses three chief questions: ``Why did we change?'' ``What are we
doing to engage students in science practices?'' and ``How did we make all of this work
online?''

Scientific truth is discovered through the practice of doing science.  \citeauthor{DBERreport}
state, ``Without [scientific] investigative practices, there would be no new scientific and
engineering knowledge.''\citep{DBERreport} Too often, our laboratory courses do not
meaningfully engage students in scientific practices, instead reproducing classical
experiments.\citep{PCAST12} This is part of a long-standing tradition of calling for authentic
engagement in scientific practices which pre-dates Lorentz transformations.\citep{AAAS1881}
Indeed, back in \citeyear{AAAS1881}, the AAAS stated our goals quite succinctly, ``[Typical
courses in t]he sciences\ldots are not made the means of cultivating the observing powers,
stimulating inquiry, exercising the judgment in weighing evidence, nor of forming independent
habits of thought.''\citep{AAAS1881}

The physics labs at our institution in the Fall of 2016 very much mirrored the national
narrative.  The lab curriculum had been largely unchanged for about 20 years, and followed the
traditional cookbook lab paradigm.  Moreover, the equipment commonly used by students was
outdated---triple beam balances and hand-drawn graphs were standard.  In the same year, as part
of an interdisciplinary team interested in engaging students in scientific practices as well as
assessing those scientific practices, we secured funding, both internally and from the NSF, to
change this status quo in physics, biology, and chemistry.  All three departments transformed
their labs using a pedagogical framework entitled, Argument Driven Inquiry
(ADI),\citep{Sampson2011,Walker2011,Walker2016} which has been specifically designed to engage
students in authentic science practices. We will focus on the physics transformation in this
paper.

We were satisfied with our progress when March 2020 rolled around, and with it, the mandatory
shut-down due to the global COVID-19 pandemic.  With students off campus, we had to abruptly
shift our delivery to an online format and quickly realized that some of the activities that we
utilized in the face-to-face sections had to be adapted to online use.  Going forward, we have
further adapted what we are doing so that students can still participate in a hands-on
laboratory experience in an online laboratory course.  We hope these changes will be meaningful
long-term as a lack of online labs is also a barrier for distance education students'
completion of a degree.  So online lab curricula would fill an institutional need.  In the next
sections, we will discuss our institutional and instructional contexts and describe each
transformation in more detail.

\section{Instructional Context}

@@

\section{Transformation \#1: Argument Driven Inquiry}
Argument Driven Inquiry (ADI) is a pedagogical method that was developed to make science labs
in school better match what actually transpires in authentic research settings
\citep{Walker2011,Sampson2011,Walker2016}.  In traditional labs, we often ask students to
answer questions, or measure quantities, that are well-known using procedures that have been
refined over time.  For example, in physics we often have students measure $g$ using a
well-known method (such as a spark timer).  In ADI labs, we give students a question to answer
(e.g., ``Does the force a fan exerts on a cart depend on the mass of the cart?''), allow them
to design a procedure for an experiment that will answer that question, collect and analyze
their data, and make claims based on their data.  We accomplish this by having students engage
in a three-part cycle:
\begin{itemize}
  \item Pre-Lab - Introducing the context
  \item Proposal Development and Data Collection
  \item Argumentation and Peer Review
\end{itemize}
In the sections below, we will comment on each part of the cycle, focusing on both students and
GTAs.

\subsection{Session 1: Pre-Lab}
Part of the goal of any laboratory is to help students develop a familiarity with certain tools
and equipment common to the lab.  The purpose of the pre-lab reading and activity is to give
students familiarity with the equipment they will be using.  For the investigation described
previously, ``Does the force a fan exerts on a cart depend on the mass of the cart?''  we need
students to become familiar with a few things, such as what we mean by a cart, and give them
some tools that they need to be able to answer this question.  As a part of the pre-lab
reading, students are given information about linear regression---both how to run it in Excel,
and how to interpret the output, and how we can estimate instantaneous velocity given position
vs.\ time data. For this experiment, we give them a cart (leaving the fan off for now) and have
them measure the acceleration of that cart down a ramp using an ultrasonic motion detector to
collect position vs.\ time data for this motion.  Students work in pairs to do this work, and
turn in a brief summary of this work on the course management system.  For this investigation,
it is the first time that they create graphs, so we set norms for graphing such as helping them
decide which variable to put on the x-axis and the y-axis as well as including units.  Pre-lab
assignments are designed to be graded in less than 1 minute/paper based on a rubric we provide
to both students and GTAs.

\subsection{Session 2: Proposal Development and Data Collection}
During this session, students work in groups of 4 to develop a proposal, get it approved by a
GTA, and then collect and analyze their data.  We have a format for proposal development that
asks students to link the scientific concept being studied---in this example case, Newton's
Second Law---to the data that they will collect.  Students also need to propose a plan for
analyzing their data and minimizing potential sources of error (such as friction).  Once
students have an approved proposal, they begin collecting and analyzing their data.  This may
lead them to refine or revise their method (which they are encouraged to do).  Grading for the
GTAs this week is minimal (full credit for completing the proposal).  However, this is the most
critical week's for GTA training.  GTAs have to be clear that the goal is to allow a diverse
number of ways of collecting/analyzing this data.  How much does the mass of the cart need to
vary?  How many different cart masses do the students need to take?  How should students deal
with friction? All of these are questions that we want the \textit{students} to answer.  And in
some sense, we will be disappointed if all of the students come up with the same answer.
Experimental methods certainly have an impact on experimental results.

\subsection{Session 3: Argumentation and Peer Review}
During this session, the students begin by sharing their results.  Students finish analyzing
their data during the prior week, and prepare a ``poster'' on a whiteboard.  One person from
the group stays at the whiteboard while the other group members (travelers) go to other groups.
The travelers learn what other groups did, and compare to their results in a structured
argumentation setting.  Groups spend about 3--5 minutes at a poster, and rotate so that they
see at least 3 other posters.  After the travelers return, critically, the groups spend time
discussing what they saw and reflecting on their own results.  What is exciting is that
sometimes, these interactions actually drive groups to change their claims \citep{Walker2019}.

Each student writes a shortened lab report consisting of three sections: an introduction to the
scientific concept, a description of the experimental method or procedure, and a discussion
section laying out their results and the evidence and justification supporting their results.
The authors have created different peer review calibration videos for the students to watch so
they can look for the same things that we are going to be grading them on when we get their
final reports.  Students complete their lab report and submit it by the end of the day.  Then,
students take part in double-confidential peer review.  We facilitate this online.  They read
and reply to two different lab reports.  Finally, they get their feedback and use this to
revise their own lab report.  At the end of this week, students turn in their revised,
peer-reviewed final lab report.  This is the heaviest week of GTA grading.  We provide a rubric
for lab report grading, and each lab report generally takes about 3-5 minutes to grade.

@@

\begin{table}
  \caption{\label{tab: 1251 fall} General Physics Laboratory I investigations for Fall 2020
    block schedule.}
  \begin{ruledtabular}
    \begin{tabular}{ccp{28em}}
      Investigation & Topic & Guiding Question\\
      \hline
      1 & 1-D kinematics & Does a ball rolling on an incline have the same acceleration on the way up as it does on the way down? \\
      2 & Periodic motion & At what nut position is the period of the physical pendulum equal to \SI{1.30}{s}? \\
      3 & 2-D collisions & Is the collision between two marbles elastic?
    \end{tabular}
  \end{ruledtabular}
\end{table}

@@

\begin{table}
  \caption{\label{tab: 1261 fall} General Physics Laboratory II investigations for Fall 2020
    block schedule.}
  \begin{ruledtabular}
    \begin{tabular}{ccp{26em}}
      Investigation & Topic & Guiding Question\\
      \hline
      1 & Current and resistance & Does a light bulb behave like a resistor? \\
      2 & Time varying circuits & Do two of the lab kit capacitors have the same capacitance? \\
      3 & Diffraction of light & Are hairs from different people the same diameter?
    \end{tabular}
  \end{ruledtabular}
\end{table}

@@

\begin{table}
  \caption{\label{tab: 1261 lab kit} General Physics Laboratory II lab kit contents.}
  \begin{tabular}{clc}
    \hline\hline
    Quantity & Item & Investigation(s)\\
    \hline
    1 & \SI{100}{\ohm} resistor & 1 \\
    1 & \SI{330}{\ohm} resistor & 1 \\
    1 &  \SI{100}{\ohm} potentiometer & 1 \\
    1 & E10 light bulb holder & 1 \\
    1 & \SI{5}{V} E10 incandescent light bulb & 1 \\
    1 & Breadboard & 1, 2, 3 \\
    1 & Breadboard power supply & 1, 2, 3 \\
    1 & USB power supply cable & 1, 2, 3 \\
    1 & Jumper wire set & 1, 2 \\
    2 & Multimeters & 1, 2 \\
    1 & Mini screwdriver for multimeters & 1,2 \\
    5 & \SI{500}{mA} fusses for multimeters & 1,2 \\
    4 & Alligator clip leads & 1, 2 \\
    1 & \SI{1}{\mega\ohm} resistor & 2 \\
    2 & \SI{100}{\micro\farad} capacitor & 2 \\
    1 & \SI{5}{V}, \SI{650}{nm} laser module & 3 \\
    1 & Tape measure with cm scale & 3 \\            
    \hline\hline
  \end{tabular}
\end{table}

\section{Conclusion}
Discussion Points:
\begin{itemize}
  \item What have been our successes?  (Related to both Transformations)
  \item What have been our challenges?
  \item What plans do we have for online labs in the future?
\end{itemize}

\begin{acknowledgments}
  Redacted for review.% This work has been supported by the NSF (Award \# 1725655).
\end{acknowledgments}

>>>>

\section{Additional instructional context}

@@

\section{Transformation \#1: Argument Driven Inquiry}
\subsection{Physics \textrm{I} Investigation 1}

\subsection{Physics \textrm{I} Investigation 2}

\subsection{Physics \textrm{I} Investigation 3}

\subsection{Physics \textrm{I} Investigation 4}

\subsection{Physics \textrm{II} Investigation 1}

\subsection{Physics \textrm{II} Investigation 2}

\subsection{Physics \textrm{II} Investigation 3}

\subsection{Physics \textrm{II} Investigation 4}

@@

\begin{table}
  \caption{\label{tab: 1251 fall} General Physics Laboratory I investigations for Fall 2020
    block schedule.}
  \begin{tabular}{ccp{28em}}
    \hline\hline
    Investigation & Topic & Guiding Question\\
    \hline
    1 & 1-D kinematics & Does a ball rolling on an incline have the same acceleration on the way up as it does on the way down? \\
    2 & Periodic motion & At what nut position is the period of the physical pendulum equal to \SI{1.30}{s}? \\
    3 & 2-D collisions & Is the collision between two marbles elastic? \\
    \hline\hline
  \end{tabular}
\end{table}

@@

\begin{table}
  \caption{\label{tab: 1261 fall} General Physics Laboratory II investigations for Fall 2020
    block schedule.}
  \begin{tabular}{ccp{26em}}
    \hline\hline
    Investigation & Topic & Guiding Question\\
    \hline
    1 & Current and resistance & Does a light bulb behave like a resistor? \\
    2 & Time varying circuits & Do two of the lab kit capacitors have the same capacitance? \\
    3 & Diffraction of light & Are hairs from different people the same diameter? \\
    \hline\hline
  \end{tabular}
\end{table}

@@

\begin{table}
  \caption{\label{tab: 1261 lab kit} General Physics Laboratory II lab kit contents.}
  \begin{tabular}{clc}
    \hline\hline
    Quantity & Item & Investigation(s)\\
    \hline
    1 & \SI{100}{\ohm} resistor & 1 \\
    1 & \SI{330}{\ohm} resistor & 1 \\
    1 & \SI{100}{\ohm} potentiometer & 1 \\
    1 & E10 light bulb holder & 1 \\
    1 & \SI{5}{V} E10 incandescent light bulb & 1 \\
    1 & Breadboard & 1, 2, 3 \\
    1 & Breadboard power supply & 1, 2, 3 \\
    1 & USB power supply cable & 1, 2, 3 \\
    1 & Jumper wire set & 1, 2 \\
    2 & Multimeters & 1, 2 \\
    1 & Mini screwdriver for multimeters & 1,2 \\
    5 & \SI{500}{mA} fuses for multimeters & 1,2 \\
    4 & Alligator clip leads & 1, 2 \\
    1 & \SI{1}{\mega\ohm} resistor & 2 \\
    2 & \SI{100}{\micro\farad} capacitor & 2 \\
    1 & \SI{5}{V}, \SI{650}{nm} laser module & 3 \\
    1 & Tape measure with cm scale & 3 \\            
    \hline\hline
  \end{tabular}
\end{table}

\clearpage
